---
output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    number_sections: false
    fig_caption: true
    toc: false
    includes:
        before_body: 
        - header.sty
geometry: "left=3cm,right=3cm,top=2cm,bottom=2cm"
linespacing: 1.5
fontsize: 11pt
mainfont: Palatino
bibliography: refs.bib
citation-style: ../csl/university-of-york-harvard.csl
link-citations: true
citecolor: teal
linkcolor: teal
urlcolor: teal
header-includes:
- \usepackage[labelsep=period]{caption}
- \usepackage[labelfont=bf]{caption}
- \usepackage[switch]{lineno}
- \usepackage{type1cm} % scalable fonts
- \usepackage{lettrine}
- \usepackage{booktabs}
- \usepackage{sectsty} \sectionfont{\centering}
- \usepackage{caption}
- \captionsetup[figure]{font=footnotesize}
- \captionsetup[table]{font=footnotesize}
- \captionsetup[table]{justification=centerlast}
- \captionsetup[figure]{justification=centerlast}
editor_options: 
    markdown:
      canonical: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load-pkgs}
library(tidyverse) # Load the tidyverse. Functions are used for tidying data and creating plots via ggplot2.

library(kableExtra) # Load kableExtra. Table generation.

library(wordcountaddin) # Load wordcountaddin. Counts words and characters in text in an R markdown document.
```

```{r source}
 source("../R/final.R", local = knitr::knit_global())
```

# Introduction

Atherosclerosis is a chronic inflammatory disease characterised by the formation of arterial plaques.
Haemodynamic shear stress has been identified as a modulator of site-specificity in atherosclerosis, which occurs preferentially in regions exposed to low, oscillatory shear stress [@stone2007].
Whereas areas of high, laminar shear stress are atheroprotective [@timmins2017].
Shear stress is an important factor in regulating gene expression in vascular endothelial cells [@Ni2010], which is thought to contribute to the susceptibility of plaque formation in atheroprone sites.
Multiple omics studies have implicated variations in flow with the regulation of developmental signalling pathways in atherosclerosis, including the Wnt Pathway [@Souilhol2020; @Gelfand2011].

Wnt is an evolutionarily conserved pathway with a critical role in axis patterning during embryonic development.
In the absence of Wnt, axin forms a destruction complex with glycogen synthase kinase 3β (GSK-3) and adenomatous polyposis coli (APC), which phosphorylates β-catenin and targets it for degradation.
However, in the active canonical Wnt pathway, Wnt ligands interact with Frizzled and LRP receptors.
This leads to the translocation of axin, inhibiting the formation of the destruction complex, allowing β-catenin to accumulate and translocate to the nucleus, where it will activate the transcription of Wnt target genes [@gordon2006].
Of these includes axin, which acts as a negative regulator of Wnt signalling [@Jho2002; @Lustig2002].

Shear stress-mediated Wnt orchestrates a range of endothelial responses, including angiogenesis, which is increased in regions of low shear stress compared to high shear stress [@du2018].
One target of Wnt, angiopoietin-2 (ANGPT2), is an established growth factor involved in angiogenesis.
Studies in both zebrafish and mice have shown that the increase in ANGPT2 contributes to the development of atherosclerosis [@Li2014-mx; @farhat2013].

Thrombospondin-1 (THBS1) is a glycoprotein involved in endothelial cell interactions and is a potential target of Wnt.
High levels of THBS1 has been correlated with the inhibition of tumour angiogenesis [@naumov2006], possibly by induction of apoptosis via the TGF-β pathway [@Miao2001; @yee2004], or by inhibition of the VEGF pathway [@gupta1999; @kaur2010].
@jo2005 demonstrated that activation of the Wnt pathway downregulates THBS1 in colon cancer.
Thus, Wnt may mediate the expression of THBS1 in response to flow.
Since atherosclerosis occurs in regions of increased angiogenesis, we would expect THSB1 to be downregulated in low shear stress.
This has been confirmed in a study by @Moura2008.

This study was conducted to investigate the RNA expression of ANGPT2 and THBS1 in HUVECs exposed to atheroprone low shear stress (LSS), compared to atheroprotective high shear stress (HSS), and whether these genes are mediated by canonical Wnt signalling.
We addressed this using an orbital shaker model, along with canonical Wnt inhibitor, XAV939.

# Materials and Methods

## Cell Culture and Application of Shear Stress

HUVECs were cultured at 37°C in 8% M199, 0.15% sodium bicarbonate, 1 U/mL pen-strep, 0.1 ug/ml amphotericin B, 20% Hi-FBS, 30 ug/ml endothelial cell growth supplement (ECGS), and 10 U/ml heparin.
After reaching \~80% confluence, passage 2 cells were incubated with 1ml of trypsin until cells thoroughly detached, and neutralised with 9ml of M199.
They were then re-suspended in M199 media before transferring to 10mm radius 6 well plates coated in 1% gelatin.
The canonical Wnt pathway was inhibited using XAV939.
Once confluent, cells were treated with either 3ml of 0.1% DMSO in M199 or 0.1% XAV939 in M199 [@Zhu2017].
They were placed on an orbital shaker at 210 rpm for 72 hours, and exposed to low (\~5 dynes/cm^2^) and high shear stress (\~15 dynes/cm^2^) [@Warboys2019], with the exception of the static control.

## RNA Extraction and Real-Time Quantitative PCR

Cells were isolated from the periphery and centre of the plates with cold PBS and centrifuged for 5 minutes at 400g.
Total mRNA was extracted using the RNEasy Mini Kit (Qiagen) and the concentration was determined spectrophotometrically.
cDNA synthesis was performed using the Verso cDNA Synthesis Kit (Thermo Scientific) as per the manufacturers instructions.
ANGPT2, AXIN2, THSB1, and HPRT1 (reference gene) mRNA was quantified using StepOne qPCR (Thermo Scientific) with SYBR Green, using oligonucleotide qPCR primers from Ensembl [@howe2020] (Table \@ref(tab:primers)).
The amplification included 30 cycles at 95°C for 30s, 5°C for 30s, and 72°C for 45s, followed by 72 °C 10 min.

```{r primers}
knitr::kable(
    data.frame(
        Gene = c(
            "ANGPT2",
            "ANGPT2",
            "AXIN1",
            "AXIN1",
            "HPRT1",
            "HPRT1",
            "THSB1",
            "THSB1"
        ),
        Direction = c("L", "R", "L", "R", "L", "R", "L", "R"),
        Sequence = c(
            "CGGCTGTGATGATAGAAATAGGGA",
            "GTTCCAAGAGCTGAAGTTCAAGTC",
            "TGTCACTTACTTTTTCTGTGGGGA",
            "TGTCACTTACTTTTTCTGTGGGGA",
            "TTGGTCAGGCAGTATAATCC",
            "GGGCATATCCTACAACAAC",
            "AAAGATGGAGAATGCTGAGTTGGA",
      "GGTTCCAAAGACAAACCTCACATT"
    )
  ),
  booktabs = T,
  caption = "Oligonucleotide qPCR primers from Ensembl."
) %>%
  collapse_rows(columns = 1, latex_hline = "none") %>%
  kable_styling(full_width = T,
                latex_options = "hold_position",
                position = "left") %>%
  column_spec(column = 3, width = "8cm")
```

## Statistical Analysis

Relative expression is expressed as 2^ΔΔCt^ fold change ± SEM, normalised to the HPRT control.
Normality was determined with Kolmogorov-Smirnov Tests.
Comparison analysis was performed using the Student's t-test.
All plots and analyses were performed in R [@R].

# Results

To asses the effect of low and high shear stress on gene express, HUVECs were exposed to flow using an orbital shaker system.
Gene expression of ANGPT2, THSB1, and Wnt reporter AXIN2 was quantified by qPCR.
Gene expression in HUVECs exposed LSS was compared to those exposed to HSS (Figure \@ref(fig:plots)A).
Low shear stress downregulated the expression of AXIN2, ANGPT2, and THBS1.
AXIN2, a known Wnt target, decreased by 0.28-fold in low shear stress.
Similarly, ANGPT2 was decreased by 0.15-fold, and THBS1 was decreased 0.12-fold.
Using the same method, HUVECs were also treated with either DMSO or canonical Wnt inhibitor XAV929, to assess for regulation by Wnt signalling.
Expression in the presence of XAV939 was then compared to DMSO (Figure \@ref(fig:plots)B).
Exposure to LSS with the addition XAV939 intensified the expression of AXIN2 by 18.82-fold, ANGPT by 33.35-fold, and THBS1 by 35.59-fold.
Whereas XAV929 decreased expression in HSS.
AXIN2 decreased 0.68-fold, ANGPT2 by 0.42-fold, and THBS1 by 0.87-fold.
Results were analysed using the Students *t* test, which regarded them insignificant due to the sample size.

```{r plots, fig.width = 8, fig.height = 4, fig.cap = "**Low shear stress downregulates gene expression in the presence Wnt, and upregulates gene expression in the absence of Wnt.**. Cells were treated with DMSO(-) or XAV939(+) and exposed to low or high shear stress. Levels of angiopoietin-2, axin-2, and thrombospondin-1 mRNA were quantified by qPCR. (**A**) Low shear stress upregulated the expression of all genes. AXIN: 0.28±0.12, ANGPT2: 0.15±0.09, THBS1: 0.12±0.15. Data is shown as fold change ± SEM of low shear stress relative to high shear stress. (**B**) XAV939 upregulated expression of all genes in low shear stress and downregulated expression of all genes in high shear stress. AXIN: 18.82±2.66 & 0.68±0.50, ANGPT2: 33.36±15.80 & 0.42±0.14, THBS1: 35.59±4.55 & 0.87±0.33. Data is shown as fold change ± SEM of XAV939 relative to DMSO."}

plots <- ggarrange(location_plot, drug_plot, nrow = 1,
    labels = c("A", "B", "C"),
    font.label = list(face = "bold", size = 16),
    label.x = -0.01,
    label.y = 1.02)
ggsave("plots.png", plots, width = 10, height = 8)
plots
```

```{=html}
<!-- 
Clear and well-structured description of findings 
Appropriate narrative of the experimental rationale and the major conclusions 
May include descriptions of problem solving and troubleshooting
Well-prepared and consistently presented tables and figures 
Numbered tables and figures with legends (captions).
Legends should provide enough information to allow the reader to understand the table/figure.
Get your figures and tables ready first
Use enough text to provide a clear narrative through those figures and tables
Do start with some basic results
Highlight the key findings, but don’t discuss them
Make sure statistics are reported in the correct format
Few references -->
```
# Discussion

In our study, we compare the effects of low shear stress and high shear stress on the expression of two regulators of angiogenesis, ANGPT2 and THBS1, and whether their expression is controlled by Wnt. This was achieved using an orbital shaker model, first described by @Warboys2014. Direct Wnt target, AXIN2, was also measured to quantify Wnt expression [@Jho2002].
Finally, we compared gene expression in LSS to HSS, and XAV939 to DMSO.

Our findings indicate that AXIN2, ANGPT, and THBS1 is downregulated in HUVECs exposed to LSS (Figure \@ref(fig:plots)A).
The downregulation of ANGPT and AXIN2 in LSS differs from our expectations.
Previous studies demonstrate that canonical Wnt is activated by low shear stress [@Gelfand2011], as is its direct target, AXIN2 [@Jho2002].
ANGPT2, a positive regulator of angiogenesis, has also been identified as a positive target of Wnt in zebrafish [@Li2014-mx].
Thus, ANGPT2 was also expected to be upregulated in LSS.
However, the results imply that Wnt is downregulated by LSS compared to HSS, and in doing so, downregulates both ANGPT2 and AXIN2.

On the other hand, the lower expression of THSB1 in low shear stress was anticipated [@Moura2008].
Wnt has been shown to inhibit THBS1, a negative regulator of angiogenesis, in colonic tumours [@jo2005].
Therefore, its expression should increase with the downregulation of AXIN2.
This was not the case, suggesting that it is not repressed by the Wnt pathway as previously reported.

\color{red}

-   This is a very different context - you should make this clear, and if it's the basis of choosing it as a target mention it in the introduction.
    If it's not regulated by Wnt in this context - then that is fine but you need to be quite explicit in how you discuss it.

\color{black}

In the presence of XAV939, all genes were upregulated in both low and high shear stress (Figure \@ref(fig:plots)B).
Initially, this would imply that Wnt is an inhibitor of ANGPT2 and THSB1.

However, since AXIN2 is a direct inhibitor of Wnt, this suggests that the inhibitor in fact failed to inhibit Wnt.
?

\color{red}

-   Maybe - how could you find out?
    Also - is this canonical or non-canonical Wnt?
    There is cross talk between the two kinds of pathway, but they are quite separate.
    XAV939 won't inhibit non-canonical Wnt signalling.

-   If Inhibitor failed, why not consistent with the DMSO samples?

    -   XAV939 works but issue with primers specificity? Bind to another protein upregulated in LSS. Wnt downregulated in HSS hence why fold changes smaller? Small sample size so downregulation seems more prominent then it actually is

\color{black}

There are many discrepancies between our results and prior studies, likely due to limitations and errors in our method...

\color{red}

-   Possible Errors and Refine Methods

    -   Small sample = not enough to confirm spectrophotometer was consistent

    -   No prior PCR experience

-   Melt curve suggests primer dimers & contamination (include images of curves or include in supplementary?)

-   en face staining for gene markers in inner curvature [@Warboys2014]

-   Direct interaction with Wnt or via VEGF?

    -   THSB1 inhibits VEGF by competing for heparin (which is used in our medium) [@gupta1999; @dias2012; @tolsma1993]

-   Methods used in other studies

-   animal studies, zebrafish, mice, pigs [@serbanovic-canic2017; @Li2014-mx; @farhat2013a; @Moura2008; @moonen2015]

-   non-canonical wnt [@franco2016]

-   relative vs quantitative expression

-   flow chamber instead of orbital shaker

\color{black}

```{=html}
<!-- Do not simply recapitulate the results
Explore the wider meaning of the results
To what extent has your investigation addressed the aims provided in the introduction?
How have your findings advanced understanding of the problem under investigation?
How do your findings relate to previous studies in this area?
Based on your findings, what new investigations might be conducted?
Work in parallel with the introduction
Don’t just give a commentary of the results
How does your work compare with previous work?
Why might you have got different results?
How could you improve the project? (Don’t just say ‘collect more data’!) -->
```
<!-- In future look at vascular repair with scratch wound assay.  Flow chamber more control over force nut expensive-->

# Acknowledgements

```{=tex}
\begin{flushright}
Word Count: `r word_count("report.Rmd") - 28` Words
\end{flushright}
```
# References

\small
